---
title: "Tidy_evaluation"
author: "Mónica Alonso"
date: "7/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(datos)
```

## Tidy evaluation: 
tipo de evaluación no estandar usada en tidyverse.  
* arrange(), count(), filter(), group_by(), mutate() y summarise() usan **data_masking** por ejemplo se pueden usar las variables como si fueran variables en el ambiente (se puede escribir mi_variable en vez de df$mi_variable).
    *env-variables: son del programa, en gral creadas con <-
    *data-variables: son variables "estáticas" que viven en el            dataframe, por ej. csv o xls), o manipulando variables existentes
    
```{r ejemplo_data_masking}
df <- data.frame(x = runif(3), y = runif(3)) #df es una env-variable
df$x   # x es una data-variable fuera de la env-variable df usando $

```
### Indirection  
Los dos problemas principales son:
 * data-variable como argumento de función, se necesitas abrazar el argumento rodeandolo con llaves dobles, como filter(df, {{ var }})
 * cuando tienes una env_variable que es un vector de caracteres, y necesitas indexar en el .data con [[,
 como ```summarise(df, mean(.data[[ var]]))```
 
```{r ejemplo_embracing}
var_summary <- function(data, var) {
  data %>% 
    summarise(n = n(), min= min({{ var }}), max = max({{ var }}))
}
mtcars %>% 
  group_by(cyl) %>% 
  var_summary(mpg)
```

```{r usando_for}
for (var in names(mtcars)) {
 mtcars %>% count(.data[[var]]) %>% print() 
}  #notar que .data no es un dataframe, es una construccion especial, 
#un pronombre, que permite acceder a la variable actual sea directamt, con .data$x indirectamente con .data[[ var ]]. No esperar otras funciones que trabajen con esto.
```
## Tidy selection  
data masking hace facil trabajar con valores de un data set. Tidy selection es una herramienta complementaria q hace facil trabajar con columnas de un data set.

Debajo de las funciones q usan tidy selection está el pkg tidyselect.
Provee un lenguaje dedominio miniatura q hace facil seleccionar columnas por nombre, posicion o tipo.
* select(df, 1) o select(df, last_col())  
* select(df, c(a, b, c)) 
* select(df, starts_with("a"))
* select(df, ends_with("z")) 
* select(df, where(is.numeric))  selecciona columnas numericas  

### Indirection    
Como en data masking, tidy selection hace faciles tareas comunes, y más duras tareas menos comunes.  Cuando quieres usar tidy select con la especificación de columna almacenada en una variable intermedia, necesitas aprender nuevas herramientas. 2 formas de indirection:
* cuando tiene data-variable en una env-variable que es 
When you have the data-variable in an env-variable that is a function argument, you use the same technique as data masking: you embrace the argument by surrounding it in doubled braces.

The following function summarises a data frame by computing the mean of all variables selected by the user:
```{r}
summarise_mean <- function(data, vars) {
  data %>% summarise(n = n(), across({{ vars }}, mean))
}
mtcars %>% 
  group_by(cyl) %>% 
  summarise_mean(where(is.numeric))
```

When you have an env-variable that is a character vector, you need to use all_of() or any_of() depending on whether you want the function to error if a variable is not found.

The following code uses all_of() to select all of the variables found in a character vector; then ! plus all_of() to select all of the variables not found in a character vector:
```{r}
vars <- c("mpg", "vs")
mtcars %>% select(all_of(vars))
mtcars %>% select(!all_of(vars))
```

How tos
The following examples solve a grab bag of common problems. We show you the minimum amount of code so that you can get the basic idea; most real problems will require more code or combining multiple techniques.

User-supplied data
If you check the documentation, you’ll see that .data never uses data masking or tidy select. That means you don’t need to do anything special in your function:

mutate_y <- function(data) {
  mutate(data, y = a + x)
}
Eliminating R CMD check NOTEs
If you’re writing a package and you have a function that uses data-variables:

my_summary_function <- function(data) {
  data %>% 
    filter(x > 0) %>% 
    group_by(grp) %>% 
    summarise(y = mean(y), n = n())
}
You’ll get an R CMD CHECK NOTE:

N  checking R code for possible problems
   my_summary_function: no visible binding for global variable ‘x’, ‘grp’, ‘y’
   Undefined global functions or variables:
     x grp y
You can eliminate this by using .data$var and importing .data from its source in the rlang package (the underlying package that implements tidy evaluation):

#' @importFrom rlang .data
my_summary_function <- function(data) {
  data %>% 
    filter(.data$x > 0) %>% 
    group_by(.data$grp) %>% 
    summarise(y = mean(.data$y), n = n())
}
One or more user-supplied expressions
If you want the user to supply an expression that’s passed onto an argument which uses data masking or tidy select, embrace the argument:

my_summarise <- function(data, group_var) {
  data %>%
    group_by({{ group_var }}) %>%
    summarise(mean = mean(mass))
}
This generalises in a straightforward way if you want to use one user-supplied expression in multiple places:

my_summarise2 <- function(data, expr) {
  data %>% summarise(
    mean = mean({{ expr }}),
    sum = sum({{ expr }}),
    n = n()
  )
}
If you want the user to provide multiple expressions, embrace each of them:

my_summarise3 <- function(data, mean_var, sd_var) {
  data %>% 
    summarise(mean = mean({{ mean_var }}), sd = mean({{ sd_var }}))
}
If you want to use the names of variables in the output, you can use glue syntax in conjunction with :=:

my_summarise4 <- function(data, expr) {
  data %>% summarise(
    "mean_{{expr}}" := mean({{ expr }}),
    "sum_{{expr}}" := sum({{ expr }}),
    "n_{{expr}}" := n()
  )
}
my_summarise5 <- function(data, mean_var, sd_var) {
  data %>% 
    summarise(
      "mean_{{mean_var}}" := mean({{ mean_var }}), 
      "sd_{{sd_var}}" := mean({{ sd_var }})
    )
}
Any number of user-supplied expressions
If you want to take an arbitrary number of user supplied expressions, use .... This is most often useful when you want to give the user full control over a single part of the pipeline, like a group_by() or a mutate().

my_summarise <- function(.data, ...) {
  .data %>%
    group_by(...) %>%
    summarise(mass = mean(mass, na.rm = TRUE), height = mean(height, na.rm = TRUE))
}

starwars %>% my_summarise(homeworld)
#> # A tibble: 49 x 3
#>   homeworld    mass height
#>   <chr>       <dbl>  <dbl>
#> 1 Alderaan       64   176.
#> 2 Aleen Minor    15    79 
#> 3 Bespin         79   175 
#> 4 Bestine IV    110   180 
#> # … with 45 more rows
starwars %>% my_summarise(sex, gender)
#> `summarise()` has grouped output by 'sex'. You can override using the `.groups` argument.
#> # A tibble: 6 x 4
#> # Groups:   sex [5]
#>   sex            gender      mass height
#>   <chr>          <chr>      <dbl>  <dbl>
#> 1 female         feminine    54.7   169.
#> 2 hermaphroditic masculine 1358     175 
#> 3 male           masculine   81.0   179.
#> 4 none           feminine   NaN      96 
#> # … with 2 more rows
When you use ... in this way, make sure that any other arguments start with . to reduce the chances of argument clashes; see https://design.tidyverse.org/dots-prefix.html for more details.

Transforming user-supplied variables
If you want the user to provide a set of data-variables that are then transformed, use across():

my_summarise <- function(data, summary_vars) {
  data %>%
    summarise(across({{ summary_vars }}, ~ mean(., na.rm = TRUE)))
}
starwars %>% 
  group_by(species) %>% 
  my_summarise(c(mass, height))
#> # A tibble: 38 x 3
#>   species   mass height
#>   <chr>    <dbl>  <dbl>
#> 1 Aleena      15     79
#> 2 Besalisk   102    198
#> 3 Cerean      82    198
#> 4 Chagrian   NaN    196
#> # … with 34 more rows
You can use this same idea for multiple sets of input data-variables:

my_summarise <- function(data, group_var, summarise_var) {
  data %>%
    group_by(across({{ group_var }})) %>% 
    summarise(across({{ summarise_var }}, mean))
}
Use the .names argument to across() to control the names of the output.

my_summarise <- function(data, group_var, summarise_var) {
  data %>%
    group_by(across({{ group_var }})) %>% 
    summarise(across({{ summarise_var }}, mean, .names = "mean_{.col}"))
}
Loop over multiple variables
If you have a character vector of variable names, and want to operate on them with a for loop, index into the special .data pronoun:

for (var in names(mtcars)) {
  mtcars %>% count(.data[[var]]) %>% print()
}
This same technique works with for loop alternatives like the base R apply() family and the purrr map() family:

mtcars %>% 
  names() %>% 
  purrr::map(~ count(mtcars, .data[[.x]]))
Use a variable from an Shiny input
Many Shiny input controls return character vectors, so you can use the same approach as above: .data[[input$var]].

library(shiny)
ui <- fluidPage(
  selectInput("var", "Variable", choices = names(diamonds)),
  tableOutput("output")
)
server <- function(input, output, session) {
  data <- reactive(filter(diamonds, .data[[input$var]] > 0))
  output$output <- renderTable(head(data()))
}
See https://mastering-shiny.org/action-tidy.html for more details and case studies.


***
1. filter() esta inspirado en subset() de R base. subset() usa data masking, pero no funciona con tidy evaluation.

1. En R, los argumentos son lazily evaluated; significa que hasta que no se usen, no tienen ningún valor. Para saber más, ver:
<https://adv-r.hadley.nz/functions.html#lazy-evaluation>  



